---
title: "Reverse Engineering Project"
author: "Student names here"
date: "Date here"
output:
  html_document:
    theme: cerulean
    highlight: pygments
    toc: true
    toc_float:
      collapsed: true
      smooth_scroll: false
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Introduction

In this notebook, we are reverse engineering the story, [STORY NAME](STORY%20URL)

## Load libraries

Loading required libraries for this analysis.

```{r echo=FALSE, message=FALSE}
library(tidyverse)
library(lubridate)
library(janitor)
```

## Load and Cleaning Data

In this section, describe the source of the data, write a basic data dictionary for data you are working with, and discuss any caveats or issues you discovered working with this data.

```{r}
# Load required data

# Path to data should be loaded from folder "data" i.e. read_csv("data/name_of_data.csv")

# Clean required data and prepare for analysis if needed. 

sports_political_donations <- read_csv("data/sports-political-donations.csv") |> clean_names()

sports_political_donations <- sports_political_donations |> mutate(amount = parse_number(amount))

#Used ChatGPT to learn about parse_number. Prompt was: how would I change data from character to a number in r. The data is for amount of money but there is a $ in front of each number, making it a character
```

## Sentences to Engineer

In this notebook, we are reverse engineering five sentences from the story.

### Sentence 1

-   **Sentence text**: Sports owners' contributions to federal elections since 2015: \$47 million (\$10m to republicans, \$1.9 million to Dems in 2020 as of publishing.
-   **Analysis summary**: Our analysis does confirm the finding, as it shows a total of \$10 million to republicans, \$1.9 million to democrats and a \$47 million total.

```{r}
# Put code to reverse engineer sentence here
 sports_political_donations |>
  summarize(total_donations = sum(amount))
 
 sports_political_donations |>
   filter(election_year == 2020) |>
   group_by(party) |>
   summarise(sum(amount))
# Display results of code below this codeblock

```

### Sentence 2

-   **Sentence text**: "That strong Republican lean is consistent with owners’ spending in the 2018 and 2016 federal elections as well. A deep search in the Federal Election Commission database of campaign finances for principal owners, controlling owners, co-owners and commissioners from the NBA, NFL, NHL, WNBA, MLB and NASCAR reveals that this deep-pocketed group has sent \$34.2 million (72.9 percent) to Republican campaigns or super PACs purely supporting Republican causes, compared to \$10.1 million (21.5 percent) to Democrats over the past three elections. Less than 6 percent of contributions went to bipartisan or unaffiliated recipients."
-   **Analysis summary**: At first, it looked like the values were slightly off from what the five thirty-eight story initially reported, but after re-doing the math, figured out that the "Bipartisan, but mostly \_\_\_\_" were included in the calculations. After re-running the numbers, everything checked out.

```{r}
# Put code to reverse engineer sentence here
sports_political_donations |> filter(str_detect(election_year, "2016|2018")) |>
  group_by(party) |>
  summarise(total = sum(amount)) |>
  mutate(percent = total/sum(total) * 100) |>
  arrange(desc(total))

# Display results of code below this codeblock

```

### Sentence 3

-   **Sentence text**: But while that overall summary is interesting, a more pronounced signal begins to emerge when we look more closely at this group’s larger political donations. There are 40 owners who have donated at least \$100,000 to Republican causes over the six-year period, compared to 23 owners topping \$100,000 to Democrats, helping lead to these splits
-   **Analysis summary**: [Write up two to three sentences describing the results of your analysis. Were you able to confirm the finding? If not, why not?]

```{r}
# Put code to reverse engineer sentence here
sports_political_donations |>
  group_by(party, owner) |>
  summarize(total_donated = sum(amount)) |>
  filter(total_donated >= 100000) |>
  summarize(count_owners = n())
# Display results of code below this codeblock

```

### Sentence 4

-   **Sentence text**: Overall, 51.7 percent of all contributions by WNBA owners went to Democrats, with 42.3 percent going to Republicans. And of the \$1.3 million contributed by WNBA owners to Republicans, 65.5 percent comes from Kelly Loeffler, the Republican from Georgia who is co-owner of the Atlanta Dream.
-   **Analysis summary**: Our analysis confirms both findings. The first section of code shows that 51.7%, or \$1.6 million, went to Democrats and 42.3%, or \$1.3 million, went to Republicans. The second section of code shows that Kelly Loeffler donated \$867,150 which accounts for 65.5% of WNBA Republican contributions.

```{r}
# Put code to reverse engineer sentence here
sports_political_donations |>
  filter(str_detect(league, "WNBA")) |>
  group_by(party) |>
  summarise(total = sum(amount)) |>
  mutate(percent = total/sum(total) * 100) |>
  arrange(desc(total))

sports_political_donations |>
  filter(str_detect(league, "WNBA") & party == "Republican") |>
  group_by(owner) |>
  summarise(total = sum(amount)) |>
  mutate(percent = total/sum(total) * 100) |>
  arrange(desc(total))
# Display results of code below this codeblock

```

### Sentence 5

-   **Sentence text**: More than 63 percent of the \$47 million in the database of nearly 3,000 contributions comes from just 108 donations split among these 40 owners — and 79 percent of that money went to Republican causes.
-   **Analysis summary**: We were able to confirm the findings that of the \$47 million donated, just over 63 percent was concentrated among 108 donations. We also found that 79 percent of that total was donated to Republican causes. 

```{r}
# Put code to reverse engineer sentence here
top_108 <- sports_political_donations |>
  arrange(desc(amount)) |>
  slice_head(n = 108)

sum(top_108$amount) / sum(sports_political_donations$amount)

top_108 |>
  filter(party == "Republican") |>
  summarise(sum(amount)) / sum(top_108$amount)
# Display results of code below this codeblock
top_108
```

-30-
