---
title: "week7_recap"
author: "Daniel Trielli"
date: "2023-10-04"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
# Turn off scientific notation
options(scipen=999)
```

## Loading the packages

Run the codeblock below to load the packages we will need for this recap

```{r}
library(tidyverse)
library(lubridate)
library(janitor)
```

## Load Data

Run the codeblock below to load the data.

```{r}
earthquakes <- read_csv('https://earthquake.usgs.gov/earthquakes/feed/v1.0/summary/all_month.csv')

#Setting time column as datetime
earthquakes <- earthquakes |> mutate(time = as_datetime(time))
```

#### Answer the questions below

Most questions have a code block and a space for an answer below. Write the code you think is necessary and, in the answer space, write out what you did and what was the result.

------------------------------------------------------------------------

#### **Q1** Do you notice anything different with this read_csv()? Particularly in the file name?

**A1: We are getting the data from a website rather than a downloaded file.**

------------------------------------------------------------------------

#### **Q2** How many records there are there in this dataset? What do they mean and what useful information we can gather from it, looking at the columns?

**A2: There are 10,825 records in this dataset showing earthquakes. We can gather information such as the highest magnitude, which locations have the most earthquakes, which earthquakes had the most depth and more.**

------------------------------------------------------------------------

#### **Q3** How do I reorganize this list to see the ones that are the deepest first? What is the depth that shows up for the deepest one, and its magnitude?

```{r}
earthquakes |>
  group_by(depth) |>
  arrange(desc(depth))
```

**A3: 750 is the deepest one and it says N/A for the magnitude.**

------------------------------------------------------------------------

#### **Q4** I just want to see the earthquakes with a magnitude larger than 6. How do I do that? And how many are there that fit this criteria?

```{r}
earthquakes |>
  filter(mag > 6)
```

**A4: There are 12 earthquakes that fit this criteria.**

------------------------------------------------------------------------

#### **Q5** What about if I want to see earthquakes that have both a magnitude larger than 6 and a depth smaller than 20? How many are there in the data set that fit [both]{.underline} these criteria?

```{r}
earthquakes |>
  filter(mag > 6 & depth < 20)
```

**A5: There are 10 earthquakes that fit both these criteria.**

------------------------------------------------------------------------

#### **Q6** What about if I want to see earthquakes that either have a magnitude larger than 6 and a depth smaller than 20? How many are there in the data set that fit [either]{.underline} these criteria?

```{r}
earthquakes |>
  filter(mag > 6 | depth < 20)
```

**A6: There are 8,614 earthquakes that fit either of these criteria.**

------------------------------------------------------------------------

#### **Q7** I'm interested in finding earthquakes that took place in Alaska. Which column should I look at? How do I use it to find all the earthquakes in Alaska? How many earthquakes took place there?

```{r}
q7<- earthquakes |>
  filter(str_detect(place, "Alaska"))
```

**A7: 3,611 earthquakes took place in Alaska.**

------------------------------------------------------------------------

#### **Q8** I notice that there is a column called 'type', that seems to have different kinds of tremors. What is the best way to find what are all the possible types of tremors, and counting how many of each there are in this data set? What are the first two most common types of tremors in this data set?

```{r}
earthquakes |>
  group_by(type) |>
  summarize(count_type = n()) |>
  arrange(desc(count_type))
```

**A8: Earthquake and quarry blast are the two most common types of tremors in this dataset.**

------------------------------------------------------------------------

#### **Q9** How do I calculate the average depth of each type of tremor? What is the average depth of the earthquake type in this data set? Is there anything that seems unusual or surprising in this finding?

```{r}
earthquakes |>
  group_by(type) |>
  summarise(mean_depth = mean(depth))
```

**A9: The average depth of the earthquake type is 22.47. It's surprising that every other type has an average depth of around 0.**

------------------------------------------------------------------------

#### **Q10** I'm interested, in the future, to see the hours in which earthquakes happen. How can I extract the hour from the time column?

```{r}
earthquakes <- mutate(earthquakes, hour = format(time, "%H"))

#Used ChatGPT for format function. Prompt was: How to extract part of POSIXCT in r
```

**A10:**

------------------------------------------------------------------------

#### **Q11** I want to make a note of all the records in this data set that I consider serious. For my purposes, I'm thinking that all tremors that have a magnitude that is larger than 3 are serious. How do I automatically create a new column making that distinction?

```{r}
earthquakes <- earthquakes |>
  mutate(serious = mag > 3)
```

**A11: Use mutate and the column will display true for serious and false for not serious.**

------------------------------------------------------------------------

#### **Q12** I have no idea how earthquakes work and I'm interested in seeing if there is a particular time of day in which serious earthquakes happen. How can I see that condensed in a table with all the hours in a day and all the serious earthquakes in each hour? What is the hour with fewer serious earthquakes and the one with the most serious earthquakes?

```{r}
earthquakes |>
  filter(serious == TRUE) |>
  group_by(hour) |>
  summarize(total_serious = sum(serious)) |>
  arrange(desc(total_serious))
```

**A12**: 6 has the most serious earthquakes and 2 has the fewest.

------------------------------------------------------------------------

#### EXTRA Q What other questions are you interested in and how would you ask them in R?

```{r}
#What is the most common hour for earthquakes in Alaska
earthquakes |>
  filter(str_detect(place, "Alaska")) |>
  group_by(hour) |>
  summarize(count_hour = n()) |>
  arrange(desc(count_hour))
```

```{r}
#How many serious earthquakes does each type have
earthquakes |>
  filter(serious == TRUE) |>
  group_by(type) |>
  summarize(total_serious = sum(serious)) |>
  arrange(desc(total_serious))
```

```{r}
#What location source had the most earthquakes
earthquakes |>
  group_by(locationSource) |>
  summarize(count_source = n()) |>
  arrange(desc(count_source))
```
